qcircuit 2.5.3 Tutorial

### 2017-11-01 CO681-QuantumInformationProcessing/assignments/assignment_03.tex
Completed 2

\begin{document}
\section{}
\subsection{a}
\paragraph{}
The phase estimation algorithm can be thought of as running a unitary matrix $W$ which maps $\ket{0}\ket{\phi}$ to $\ket{a}\ket{\phi}$ for any eigenvector $\ket{\phi}$ of $U$ with eigenvalue $\omega^a$, and then measuring the first register. So if we consider $\alpha\ket{\psi} + \beta\ket{\phi}$ then running phase estimation yields
$$W(\ket{0}(\alpha\ket{\psi} + \beta\ket{\phi})) = \alpha W(\ket{0}\ket{\psi}) + \beta W(\ket{0}\ket{\phi}) = \alpha \ket{a}\ket{\psi} + \beta\ket{b}\ket{\phi}$$
before measuring. So measuring the first register yields $\ket{a}$ with probability $|\alpha|^2$ and $b$ with probability $|\beta|^2$.
\subsection{b}
\paragraph{}
Let $\ket{\psi}$ be a normal vector in $\C^k$. Let $\ket{\phi^i}$ for $i=1,\dots, k$ be orthonormal eigenvectors of $U$ with corresponding eigenvalues $\lambda_1, \dots, \lambda_k$. Then since the set of $\ket{\phi^i}$ is a basis, there exist $\alpha_1, \dots, \alpha_k$ such that
$$\ket{\psi} = \sum_{i=1}^k \alpha_i \ket{\phi^i}.$$
So then $W$ acts on $\ket{\psi}$ as
$$W\ket{\psi} = \sum_{i=1}^k \alpha_i \ket{\lambda_i}\ket{\phi^i}.$$
Measuring the second register then yields $\ket{\phi^i}$ with probability $|\alpha_i|^2$. So then if we let $A_i = \ket{\phi^i}\bra{\phi^i}$ then $A_1, \dots, A_k$ forms a general measurement. If we let $\rho = \ket{\psi}\bra{\psi}$ then $Tr(A_i\rho A_i^\dagger) = |\alpha|^2$ and $\frac{1}{Tr(A_i\rho A_i^\dagger)} A_i \rho A_i^\dagger = \ket{\phi^i}\bra{\phi^i}$ as desired.
\section{}
\subsection{a}
\paragraph{}
The probability of getting outcome $j$ is
$$Tr(A_j\ket{\psi_j}\bra{\psi_j}A_j^\dagger) = \frac{1}{2} Tr(\ket{\psi_j}\bra{\psi_j}\ket{\psi_j}\bra{\psi_j}\ket{\psi_j}\bra{\psi_j}) = \frac{1}{2}Tr(\ip{\psi}{\psi}^3) = \frac{1}{2}.$$
\subsection{b}
\paragraph{}
Let $j \in \{0,1,2,3\}$. The density matrix resulting from measuring $\ket{\psi_j}$ is
$$\frac{1}{2}\ket{\psi_j}\bra{\psi_j} + \frac{1}{2}\ip{\psi_j}{\psi_{j+1}}^2\ket{\psi_{j+1}}\bra{\psi_{j+1}} + \frac{1}{2}\ip{\psi_j}{\psi_{j+3}}^2\ket{\psi_{j+3}}\bra{\psi_{j+3}}$$
where addition in the subscripts is taken modulo $4$. This simplifies as
$$\frac{1}{2}\ket{\psi_j}\bra{\psi_j} + \frac{1}{4}\ket{\psi_{j+1}}\bra{\psi_{j+1}} + \frac{1}{4}\ket{\psi_{j+3}}\bra{\psi_{j+3}}.$$
\end{document}

>>>>

\begin{document}
\section{}
\subsection{a}
\paragraph{}
The phase estimation algorithm can be thought of as running a unitary matrix $W$ which maps $\ket{0}\ket{\phi}$ to $\ket{a}\ket{\phi}$ for any eigenvector $\ket{\phi}$ of $U$ with eigenvalue $\omega^a$, and then measuring the first register. So if we consider $\alpha\ket{\psi} + \beta\ket{\phi}$ then running phase estimation yields
$$W(\ket{0}(\alpha\ket{\psi} + \beta\ket{\phi})) = \alpha W(\ket{0}\ket{\psi}) + \beta W(\ket{0}\ket{\phi}) = \alpha \ket{a}\ket{\psi} + \beta\ket{b}\ket{\phi}$$
before measuring. So measuring the first register yields $\ket{a}$ with probability $|\alpha|^2$ and $b$ with probability $|\beta|^2$.
\subsection{b}
\paragraph{}
Let $\ket{\psi}$ be a normal vector in $\C^k$. Let $\ket{\phi^i}$ for $i=1,\dots, k$ be orthonormal eigenvectors of $U$ with corresponding eigenvalues $\lambda_1, \dots, \lambda_k$. Then since the set of $\ket{\phi^i}$ is a basis, there exist $\alpha_1, \dots, \alpha_k$ such that
$$\ket{\psi} = \sum_{i=1}^k \alpha_i \ket{\phi^i}.$$
So then $W$ acts on $\ket{\psi}$ as
$$W\ket{\psi} = \sum_{i=1}^k \alpha_i \ket{\lambda_i}\ket{\phi^i}.$$
Measuring the second register then yields $\ket{\phi^i}$ with probability $|\alpha_i|^2$. So then if we let $A_i = \ket{\phi^i}\bra{\phi^i}$ then $A_1, \dots, A_k$ forms a general measurement since the set $A_i$'s is a set of $k$ rank $1$ orthogonal projectors. If we let $\rho = \ket{\psi}\bra{\psi}$ then $Tr(A_i\rho A_i^\dagger) = |\alpha|^2$ and $\frac{1}{Tr(A_i\rho A_i^\dagger)} A_i \rho A_i^\dagger = \ket{\phi^i}\bra{\phi^i}$ as desired.
\section{}
\subsection{a}
\paragraph{}
The probability of getting outcome $j$ is
$$Tr(A_j\ket{\psi_j}\bra{\psi_j}A_j^\dagger) = \frac{1}{2} Tr(\ket{\psi_j}\bra{\psi_j}\ket{\psi_j}\bra{\psi_j}\ket{\psi_j}\bra{\psi_j}) = \frac{1}{2}Tr(\ip{\psi}{\psi}^3) = \frac{1}{2}.$$
\subsection{b}
\paragraph{}
Let $j \in \{0,1,2,3\}$. The density matrix resulting from measuring $\ket{\psi_j}$ is
$$\frac{1}{2}\ket{\psi_j}\bra{\psi_j} + \frac{1}{2}\ip{\psi_j}{\psi_{j+1}}^2\ket{\psi_{j+1}}\bra{\psi_{j+1}} + \frac{1}{2}\ip{\psi_j}{\psi_{j+3}}^2\ket{\psi_{j+3}}\bra{\psi_{j+3}}$$
where addition in the subscripts is taken modulo $4$. This simplifies as
$$\frac{1}{2}\ket{\psi_j}\bra{\psi_j} + \frac{1}{4}\ket{\psi_{j+1}}\bra{\psi_{j+1}} + \frac{1}{4}\ket{\psi_{j+3}}\bra{\psi_{j+3}}.$$
\subsection{c}
We implement the desired unitary matrix $U$ via the following quantum circuit
\[ \Qcircuit @C=1em @R=1em {
 \ket{k}\ \  \ & \qw &  \gate{H} & \qw\\
  \ket{0}\ \ \  & \gate{H} & \ctrl{-1} & \qw
}\]
with $k \in \{0,1\}$. When $k=0$ the circuit outputs the state $$\frac{1}{\sqrt{2}}\ket{00} + \frac{1}{2}\ket{01} + \frac{1}{2}\ket{11}$$ and when $k=1$ then circuit outputs the state $$\frac{1}{\sqrt{2}}\ket{10} + \frac{1}{2}\ket{01} - \frac{1}{2}\ket{11}$$
as desired.
\end{document}
